---
title: '20160607'
author: "Meilin Yan"
date: "June 7, 2016"
output: pdf_document
---

## LOAD PACKAGES
```{r message=FALSE, warning=FALSE}
library(lubridate)
library(dplyr)
library(dlnm)
library(splines)
library(mvmeta)
```

## DATA PREP
### Read health data
```{r}
readCity <- function(city = c(), collapseAge = TRUE){
  root <- "~/tmp/NMMAPS/"
  file <- paste(root, city, ".rds", sep = "")
  df <- readRDS(file)

  if(collapseAge == TRUE){
    df <- rowsum(df[,c("alldeath", "death", "accident", "cvd", "resp",
                       "copd", "suicide", "tmpd", "dptp")], df$date,
                 reorder = FALSE)
    df$date <- as.Date(rownames(df))
    df[ , c("tmpd", "dptp")] <- df[ , c("tmpd", "dptp")]/3
    df$dow <- as.factor(as.POSIXlt(df$date)$wday)
  }
  
  return(df)
}
# example
# miam <- readCity(city = "miam", collapseAge = TRUE)
```

### Read storm data
```{r}
# `criterion` can take the values "rain125", "rain100", "rain75", "rain50"
readStorm <- function(rain = c(), city = c()){
  root <- "exposure/"
  file <- paste(root, rain, "/", city, ".rds", sep = "")
  df <- readRDS(file)
  df <- as.data.frame(df)
  df$date <- lubridate::ymd(df$closest_date)
  
  return(df)
}
# example
# miam_75 <- readStorm(rain = "rain75", city = "miam")
```

### Combine health data and storm data
```{r}
city.storm <- function(rain = c(), city = c()){
  # Get health data
  h.df <- readCity(city)
  # Get storm data
  s.df <- readStorm(rain, city)
  
  storm_days <- s.df$date
  
  df <- left_join(h.df, s.df, by = "date") %>%
        mutate(hurr = ifelse(date %in% storm_days, 1, 0))
  df$hurr <- as.factor(df$hurr)
  
  # Add abbreviate and full cityname
  df$city <- city
  sub.county <- readRDS("data/sub.county.rds")
     
  cityname <- sub.county$citynameU[sub.county$city == city]
  df$cityname <- cityname
  
  #df$time <- scale(as.numeric(df$date), center = TRUE, scale = FALSE)
  #n.years <- length(unique(as.POSIXlt(df$date)$year))
  df$year <- as.factor(year(df$date))
  df$doy <- yday(df$date)
  
  # Total death
  df$all <- df$accident + df$death
  
  return(df)
}
# example
# miam_df75 <- city.storm(rain = "rain75", city = "miam")
```

## Case-crossover
```{r}
cross_mod <- function(rain = c(), city = c(), cause = "all", control.ratio = 15){
  df <- city.storm(rain, city)
  
  case_dates <- subset(df, hurr == 1) # hurr is exposure
  control_dates <- subset(df, hurr == 0)
  
  for(i in 1:nrow(case_dates)){
  control_range <- case_dates$doy[i] + c(-3, -2, 2, 3) # exclude three-day window hurr
  control_subset <- subset(control_dates, 
                           control_dates$year != case_dates$year[i] &
                           doy %in% control_range)
  controls <- sample_n(control_subset, control.ratio)
  stratum <- paste("stratum", i, sep = ".")
  i_stratum <- rbind(case_dates[i, ], controls)
  i_stratum$stratum <- stratum

  if(i == 1){
    new_df <- i_stratum
  } else {
    new_df <- rbind(new_df, i_stratum)
  }
}
  new_df$stratum <- factor(new_df$stratum)
  # merge with df
  new_df <- select(new_df, date, stratum)
  df <- left_join(df, new_df, by = "date")
  
  # crossbasis
  cb <- crossbasis(df$hurr, lag = c(0, 13), 
                   argvar = list(fun = "lin"),
                   arglag = list(fun = "ns", knots = c(1, 4, 8)))
  
  df$year <- as.numeric(df$year)
  mod <- glm(df[, cause] ~ cb + stratum + ns(year, 2), 
             family = quasipoisson(link = log),
             data = df,
             control = glm.control(epsilon = 10E-8, maxit = 5000))
  # crosspred
  pred <- crosspred(basis = cb, model = mod, cen = 0)
  city_mod <- list(mod = mod, pred = pred)
  return(city_mod)
}

# example
a <- cross_mod(rain = "rain75", city = "miam")
```

[MY: In the above chunk, the data used for the model is not a time-series data set. So it seems I can't use crossbasis for glm. I think we may still need to do time-stratified (e.g year:month:dow) if we descide to use a case-crossover design and hope to show the lag curve.]

## Intercept only meta-regression
### First-stage: Single lag distribution of linear hurricant exposure

```{r}
model_city <- function(rain = "rain75", city, cause = "all"){
  df <- city.storm(rain, city)
  cb <- crossbasis(df$hurr, lag = c(0, 13), 
                     argvar = list(fun = "lin"),
                     arglag = list(fun = "ns", knots = c(1, 4, 8)))
  mod_results <- glm(df[ , cause] ~ cb + ns(time, 7 * 19) + factor(dow), 
             family = quasipoisson(link = log),
             data = df,
             control = glm.control(epsilon = 10E-8, maxit = 5000))
  city_pred <- crosspred(basis = cb, model = mod_results, cen = 0)
  city_mod <- list(mod_results = mod_results, city_pred = city_pred)
  return(city_mod)
}
# example
# miam_mod <- model_city(rain = "rain75", city = "miam", cause = "all")
# coef(miam_mod[["city_pred"]])
# vcov(miam_mod[["city_pred"]])
```

```{r}
plot_city_mod <- function(city_mod, city_name){
  plot(city_mod[["city_pred"]], xlab = "Lag",
       ylab = "RR compared to non-storm day",
       main = city_name, exp = TRUE, ptype = "slices", var = 1)
}
# example
# miam_mod <- model_city(rain = "rain75", city = "miam", cause = "all")
# plot_city_mod(miam_mod, city_name = "Miami, FL")
```


```{r echo = FALSE, fig.width = 12, fig.height = 10}
this_plot <- function(city, city_name, cause = "all"){
  city_mod <- model_city(rain = "rain75", city = city, cause = cause)
  plot_city_mod(city_mod, city_name)
}

oldpar <- par(mfrow = c(2, 3))
this_plot("miam", "Miami, FL")
this_plot("jckv", "Jacksonville, FL")
this_plot("no", "New Orleans, LA")
this_plot("mobi", "Mobile, AL")
this_plot("lkch", "Lake Charles, LA")
this_plot("new", "Newport News, VA")
par(oldpar)
```

```{r echo = FALSE, fig.width = 12, fig.height = 10}
oldpar <- par(mfrow = c(2, 3))
this_plot("miam", "Miami, FL", cause = "accident")
this_plot("jckv", "Jacksonville, FL", cause = "accident")
this_plot("no", "New Orleans, LA", cause = "accident")
this_plot("mobi", "Mobile, AL", cause = "accident")
this_plot("lkch", "Lake Charles, LA", cause = "accident")
this_plot("new", "Newport News, VA", cause = "accident")
par(oldpar)
```

```{r}
read_and_fit <- function(rain = "rain75", city, cause = "all"){
  print(city)
  city_mod <- model_city(rain, city = city, cause = cause)
  return(city_mod)
}

pull_city_coef <- function(city_mod){
  city_coef <- coef(city_mod[["city_pred"]])
  return(city_coef)
}

pull_city_vcov <- function(city_mod){
  city_vcov <- vcov(city_mod[["city_pred"]])
  return(city_vcov)
}

meta_cities <- function(city_list = c("miam", "no", "jckv", "mobi", 
                                      "lkch", "new"), 
                        rain = "rain75", cause = "all"){
  city_models <- lapply(city_list, read_and_fit, rain = rain, cause = cause)
  
  city_coefs <- lapply(city_models, pull_city_coef)
  city_coefs <- do.call("rbind", city_coefs)
  
  city_vcovs <- lapply(city_models, pull_city_vcov)
  
  meta_model <- mvmeta::mvmeta(city_coefs ~ 1, S = city_vcovs, method = "reml")
  return(meta_model)
}

# example
# ex_meta <- meta_cities()

pred_meta <- function(meta_model){
  hurr_basis <- dlnm::crossbasis(x = c(rep(0, 20), 1, rep(0, 20)), 
                                 lag = c(0, 13),
                                 argvar = list(fun = "lin"),
                                 arglag = list(fun = "ns", knots = c(1, 4, 8)))
  meta_pred <- crosspred(hurr_basis, coef = coef(meta_model),
                         vcov = vcov(meta_model), model.link = "log",
                         cen = 0, at = 1)
  return(meta_pred)
}

# example
# ex_meta <- meta_cities()
# ex_pred <- pred_meta(ex_meta)

plot_meta <- function(meta_pred){
  plot(meta_pred, xlab = "Lag", 
       ylab = "RR compared to non-storm day",
       exp = TRUE, ptype = "slices", var = 1)
}

# example
# ex_meta <- meta_cities()
# ex_pred <- pred_meta(ex_meta)
# plot_meta(ex_pred)

# another example
# ex_meta <- meta_cities(cause = "accident")
# ex_pred <- pred_meta(ex_meta)
# plot_meta(ex_pred)
```

```{r}
this_plot <- function(city_list, rain = "rain75", cause = "all"){
  meta_model <- meta_cities(city_list = city_list)
  meta_pred <- pred_meta(meta_model = meta_model)
  plot_meta(meta_pred = meta_pred)
}

rain_75_cities <- gsub(".rds", "", list.files("exposure/rain75"))
this_plot(city_list = rain_75_cities, rain = "rain75", cause = "all")

rain_75_cities <- gsub(".rds", "", list.files("exposure/rain75"))
this_plot(city_list = rain_75_cities, rain = "rain75", cause = "accident")

gulf_coast_cities <- c("stpe", "tamp",
                       "mobi", "no", "batr", "lafy", "lkch",
                       "hous", "corp")
this_plot(city_list = gulf_coast_cities, rain = "rain75", cause = "all")

atlantic_coast_cities <- c("miam", "jckv", "orla", "nor", "new",
                           "nwk", "jers", "ny")
this_plot(city_list = atlantic_coast_cities, rain = "rain75", cause = "all")

```


### Seocond-stage: Meta-regression by using "reml" method

The following function "Rain.glm.meta" will run the first-stage (location-specific regression) and the second-stage (meta-regression), and return the meta-regression model.

```{r}
Rain.glm.meta <- function(rain = c(), cause = c()){
  # Read the combination dataset 
  root <- "data/cities."
  file <- paste(root, rain, ".rds", sep = "")
  df <- readRDS(file)
  
  # Create a list with the location series
  locations <- unique(df$city) 
  data <- lapply(locations, function(x) df[df$city == x, ])
  names(data) <- locations
  
  # Create matrix/list to store the cumulative coefs/vcov
  y <- matrix(NA, length(data), 1, # one coefficient for city-specific analysis
                  dimnames=list(locations, paste("b", seq(1), sep="")))
  S <- vector("list", length(data))
  names(S) <- locations

  # Run first-stage model and store results
for(i in seq(data)){
  sub.data <- data[[i]]
  
  suppressWarnings(
    cb <- crossbasis(sub.data$hurr, lag = c(0, 10), 
                     argvar = list("lin", cen = 0),
                     arglag = list("integer"))
  )
  fit <- glm(sub.data[, cause] ~ cb + ns(dptp, 3) + ns(tmpd, 3) + strata , 
             family = quasipoisson(link = log),
             data = sub.data,
             control = glm.control(epsilon = 10E-8, maxit = 10000))
  
  cr <- crossreduce(cb, fit)
  
  # Store the overall cumulative coefficients and (co)variance
  y[i, ] <- cr$coefficients
  S[[i]] <- cr$vcov
}
  # Run second-stage model (univariate meta-regression)
  meta.fit <- mvmeta(y~1, S, method = "reml")
  return(meta.fit)
  
  # Add meta variable: latitude, population
#  city.char <- readRDS("data/city.char.rds")
#  pop <- city.char$pop[city.char$city %in% locations]
#  lat <- city.char$lat[city.char$city %in% locations]
  
  # Multivariate meta-regression
#  meta.pop <- update(meta.fit, . ~ pop)
#  meta.lat <- update(meta.fit, . ~ lat)
#  return(list("meta.fit" = meta.fit, "meta.pop" = meta.pop, "meta.lat" = meta.lat))
}
```

Extract point estimates and 95%CI from meta-regression
```{r}
Rain.ests <- as.data.frame(matrix(NA, 3, 7))
colnames(Rain.ests) <- c("rain", paste0("all", ".", c("point", "low", "high"), sep = ""),
                                 paste0("acci", ".", c("point", "low", "high"), sep = ""))
t.rain <- c("rain75", "rain125", "rain175")
rownames(Rain.ests) <- t.rain
Rain.ests$rain <- as.factor(c("1-rain75", "2-rain125", "3-rain175"))

for(rain in t.rain){
  # All death
  meta.all <- summary(Rain.glm.meta(rain, "all"))
  coef.all <- meta.all[[1]]
  Rain.ests[rain, 2:4] <- coef.all[1, c(1, 5, 6)]
  
  # Accident death
  meta.acci <- summary(Rain.glm.meta(rain, "accident"))
  coef.acci <- meta.acci[[1]]
  Rain.ests[rain, 5:7] <- coef.acci[1, c(1, 5, 6)]
}
# It takes couple of mins to run. So I save it for convenience.
saveRDS(Rain.ests, file = "data/Rain.ests.rds")
# Rain.ests is used to plot.
```

```{r}
Rain.ests <- readRDS("data/Rain.ests.rds")
pdf("Plots/Pooled_estimates_by_rain .pdf", width = 12, height = 5)
p1 <- ggplot(Rain.ests, aes(rain, all.point)) +
         geom_pointrange(aes(ymin = all.low, ymax = all.high)) + 
         scale_y_continuous("Pooled estimates for all-cause death", 
                            limits = c(-6.6, 6.6)) +
         scale_x_discrete("Rain Limit", labels = c("75mm", "125mm", "175mm"))

p2 <- ggplot(Rain.ests, aes(rain, acci.point)) +
         geom_pointrange(aes(ymin = acci.low, ymax = acci.high)) +
         scale_y_continuous("Pooled estimates for accidental death", 
                            limits = c(-6.6, 6.6)) +
         scale_x_discrete("Rain Limit", labels = c("75mm", "125mm", "175mm"))

grid.arrange(p1, p2, ncol = 2)
dev.off()
```
