---
title: "Summary_Analysis_1"
author: "Meilin Yan"
date: "June 9, 2016"
output: pdf_document
---
## Load packages
```{r message=FALSE, warning=FALSE}
library(hurricaneexposure)
library(hurricaneexposuredata)
library(XML)
library(dplyr)
library(ggplot2)
library(knitr)
library(choroplethr)
library(tools)
library(ggthemes)
```

Run this code if you don't already have `hurricaneexposuredata` installed (also, you'll want to reinstall the newest version of `hurricaneexposure`):

```{r eval = FALSE}
library(devtools)
install_github("geanders/hurricaneexposuredata")
install_github("geanders/hurricaneexposure")
```

## Data preparation
### Get county data of NMMAPS
```{r}
url <- "http://www.ihapss.jhsph.edu/data/NMMAPS/documentation/counties.htm"
county <- readHTMLTable(url, header = T, stringsAsFactors = F)
county <- county[[2]]

# Change fips for Miami
county$fips[county$fips == "12025"] <- "12086"

# Add 0 to these 4-digit fips
county$fips <- sprintf("%05s", county$fips)

# Save county
saveRDS(county, file = "data/county.rds")

# Save unique city&cityname
sub.county <- select(county, city, citynameU, state) %>%
                 distinct()
saveRDS(sub.county, file = "data/sub.county.rds")
```

### Combine with storm data
```{r}
county <- readRDS("data/county.rds")
communities <- subset(county, select = c("city", 
                                         "cityname", "fips"))
colnames(communities)[1] <- "commun"
```

#### Create data sets based on rain_exposure
```{r eval=FALSE}
# 50mm rain, 500km dist, 3-day window
rain_exposure(locations = communities,
              start_year = 1988, end_year = 2005,
              rain_limit = 50, dist_limit = 500,
              #days_included = c(-1, 0, 1),
              out_type = "rds",
              out_dir = "exposure/rain50")

# 75mm rain, 500km dist, 3-day window
rain_exposure(locations = communities,
              start_year = 1988, end_year = 2005,
              rain_limit = 75, dist_limit = 500,
              #days_included = c(-1, 0, 1),
              out_type = "rds",
              out_dir = "exposure/rain75")

# 100mm rain, 500km dist, 3-day window
rain_exposure(locations = communities,
              start_year = 1988, end_year = 2005,
              rain_limit = 100, dist_limit = 500,
              #days_included = c(-1, 0, 1),
              out_type = "rds",
              out_dir = "exposure/rain100")

# 125mm rain, 500km dist, 3-day window
rain_exposure(locations = communities,
              start_year = 1988, end_year = 2005,
              rain_limit = 125, dist_limit = 500,
              #days_included = c(-1, 0, 1),
              out_type = "rds",
              out_dir = "exposure/rain125")
```

#### Create data sets based on wind_exposure
```{r eval=FALSE}
# wind_limit = 12m/sec
wind_exposure(locations = communities,
              start_year = 1988, end_year = 2005, wind_limit = 12,
              out_type = "rds", out_dir = "exposure/wind12")

# wind_limit = 15m/sec
wind_exposure(locations = communities,
              start_year = 1988, end_year = 2005, wind_limit = 15,
              out_type = "rds", out_dir = "exposure/wind15")

# wind_limit = 18m/sec
wind_exposure(locations = communities,
              start_year = 1988, end_year = 2005, wind_limit = 18,
              out_type = "rds", out_dir = "exposure/wind18")
```

#### Create data sets based on distance_exposure
```{r eval=FALSE}
# dist_limit = 50km
distance_exposure(locations = communities,
                  start_year = 1988, end_year = 2005, dist_limit = 50,
                  out_type = "rds", out_dir = "exposure/dist50")
# dist_limt = 75km
distance_exposure(locations = communities,
                  start_year = 1988, end_year = 2005, dist_limit = 75,
                  out_type = "rds", out_dir = "exposure/dist75")
# dist_limit = 100km
distance_exposure(locations = communities,
                  start_year = 1988, end_year = 2005, dist_limit = 100,
                  out_type = "rds", out_dir = "exposure/dist100")
```

## Combine and summarise
### Combine function
rain_exposure function yields data sets by city, I am going to combine these data sets into one with comb.city function. 
```{r}
# 'thres' can take the value "rain50", "rain75", "rain100", "rain125","wind12", "wind15", "wind18"
comb.city <- function(thres = c()){
  root <- paste("exposure/", thres, sep = "")
  city.name <- list.files(root, full.names = TRUE)
  df.total <- data.frame()
  for(name in city.name){
    do <- readRDS(name)
    df <- as.data.frame(do)
    df.total <- rbind(df.total, df)}
    return(df.total)
  }
```

### Count No. of cities and storms
```{r}
count <- function(df = c()){
  cities <- unique(df$loc)
  No.c <- length(cities)
  storms <- unique(df$storm_id)
  No.s <- length(storms)
  return(c(No.c, No.s))
}
# rain
df50 <- comb.city("rain50")
df75 <- comb.city("rain75")
df100 <- comb.city("rain100")
df125 <- comb.city("rain125")

# wind
df12 <- comb.city("wind12")
df15 <- comb.city("wind15")
df18 <- comb.city("wind18")

# Statistics summary
No.of.cities <- c(count(df50)[1], count(df75)[1], count(df100)[1], count(df125)[1])
No.of.storms <- c(count(df50)[2], count(df75)[2], count(df100)[2], count(df125)[2])
ta <- cbind(No.of.cities, No.of.storms)
rownames(ta) <- c("50mm", "75mm", "100mm", "125mm")
kable(ta)
```

## Plots
### Plot cumulative rain of each storm by city
```{r}
# get data sets
plot_rain <- df50

# add city fullname
# sub.county <- readRDS("data/sub.county.rds")
city.abbr <- plot_rain$loc
for(city in city.abbr){
    plot_rain$cityname[plot_rain$loc == city] <-
      paste(sub.county$citynameU[sub.county$city == city],
            sub.county$state[sub.county$city == city],
            sep = ", ")
  }

# group by maximum of max_rain 
plot_rain <- dplyr::group_by(plot_rain, loc) %>%
    dplyr::mutate(loc_max_rain = max(max_rain)) %>%
    ungroup() %>%
    arrange(loc_max_rain) %>%
    mutate(cityname = factor(cityname, levels = unique(cityname)))

# categorize max_rain for plotting
plot_rain$rain_cate <- cut(plot_rain$max_rain, 
                           breaks = c(50, 75, 100, 125, Inf),
                           labels = c(1:4))  
# plot
ggplot(plot_rain, aes(max_rain, cityname, group = cityname)) +
         geom_vline(xintercept = c(50, 75, 100, 125), colour = "grey", linetype = "longdash") +
         geom_line(aes(y = cityname), colour = "grey45") +
         geom_point(aes(color = rain_cate), alpha = 0.8, size = 1.8) +
         ylab("") +
         xlab("Cumulative rain (mm)") +
         scale_x_continuous(breaks = c(50, 75, 100, 125, 200, 300),
                            limits = c(45, 300)) +
         theme_few() + 
         theme(legend.position = "none") +
         ggsave("Plots/rain_expo.png", width = 11, height = 15)
```

[BA: You need to increase you x-limit, because you're chopping off the worst storms in New Orleans and Lake Charles. I've increased it to 300 mm in the plot.]

[BA: We don't need to use color here to show the order of intensity, because we're already showing that with the position. Instead, we could consider using color to show something like region or population. We can think about what might be most meaningful.]

### Plot wind speed of each storm by city
```{r}
plot_wind <- df12

# add city fullname
# sub.county <- readRDS("data/sub.county.rds")
city.abbr <- plot_wind$loc
for(city in city.abbr){
    plot_wind$cityname[plot_wind$loc == city] <-
      paste(sub.county$citynameU[sub.county$city == city],
            sub.county$state[sub.county$city == city],
            sep = ", ")
  }

# group by maximum of max_wind 
plot_wind <- dplyr::group_by(plot_wind, loc) %>%
    dplyr::mutate(loc_max_wind = max(max_wind)) %>%
    ungroup() %>%
    arrange(loc_max_wind) %>%
    mutate(cityname = factor(cityname, levels = unique(cityname)))

# categorize max_rain for plotting
plot_wind$wind_cate <- cut(plot_wind$max_wind, 
                           breaks = c(12, 15, 18, Inf),
                           labels = c(1:3))  
# plot
ggplot(plot_wind, aes(max_wind, cityname, group = cityname)) +
         geom_vline(xintercept = c(12, 15, 18), colour = "grey", linetype = "longdash") +
         geom_line(aes(y = cityname), colour = "grey45") +
         geom_point(aes(color = wind_cate), alpha = 0.8, size = 1.8) +
         ylab("") +
         xlab("Wind speed(m/sec)") +
         scale_x_continuous(breaks = c(12, 15, 18, 25, 35, 45, 55),
                            limits = c(10, 55)) +
         theme_few() + 
         theme(legend.position = "none") +
         ggsave("Plots/wind_expo.png", width = 11, height = 15)
```


## Select the 5 worst storms based on rain/wind exposure metrics
```{r}
df50 <- arrange(df50, desc(max_rain))
head(df50)
# The five worst storms are Isidore-2002 in New Orleans(no), Allison-1989 in Lake Charles(lkch), Irene-1999 in Miami(miam), Frances-2004 in Jacksonville(jckv), and Isidore-2002 in Mobile(mobi).

df12 <- arrange(df12, desc(max_wind))
head(df12)
# The five worst storms are Andrew-1992 in Miami(miam), Katrina-2005 in New Orleans(no), Ivan-2004 and Georges-1998 in Mobile(mobi), and Andrew-1992 in Lafayette(lafy).
```


## Map
Prepare data set for mapping
```{r warning=FALSE, message=FALSE}
data("df_pop_county")
df_county <- df_pop_county 
#  mutate(value = 0)

# Add 0 to regions with 4-digit fips
df_county$region <- sprintf("%05s", df_county$region)
```

Get all the counties in the NMMAPS with storm based on 50 rain, 500 dist
```{r}
fips <- unique(county$fips)
rain_exposure(locations = fips,
              start_year = 1988, end_year = 2005,
              rain_limit = 50, dist_limit = 500,
              out_type = "rds",
              out_dir = "data/county_w_storm_rain50")
co <- list.files("data/county_w_storm_rain50")
county_w_storm <- tools::file_path_sans_ext(co)
# 75 counties
```

Set value for counties in NMMAPS
```{r}
# value=1 for county in NMMAPSE, value=2 for county in NMMAPS with storm
get.city.county <- mutate(county, value = ifelse(fips %in% county_w_storm, 2, 1)) %>% 
                   select(city, fips, value)

# For multiple counties belong to one city, they were all exposed to storm if one of them had storm.
get.city.county$value[get.city.county$city == "dlft"] <- 2
get.city.county$value[get.city.county$city == "kan"] <- 2
colnames(get.city.county) <- c("city", "region", "value")
```

Combine 
```{r}
df_county <- left_join(df_county, get.city.county, by = c("region")) %>%
             select(region, value.y)
colnames(df_county) <- c("region", "value")
df_county$value[is.na(df_county$value)] <- 0
df_county$value <- as.factor(df_county$value)
df_county$region <- as.numeric(df_county$region)
```

### Map county

The map correctly shows all the counties in NMMAPS with/without storm. However, I still couldn't figure out how to add a circle to all the NMMAPS couties. The choropleth map seems different from a common map even though I can get the reference map for a choropleth. I am trying to figure it out. 

```{r}
eastern_states <- c("alabama", "arkansas", "connecticut",
                    "delaware", "district of columbia", "florida",
                    "georgia", "illinois", "indiana", "iowa", 
                    "kansas", "kentucky", "louisiana", "maine",
                    "maryland", "massachusetts", "michigan",
                    "mississippi", "missouri", "new hampshire", 
                    "new jersey", "new york", "north carolina", 
                    "ohio", "oklahoma", "pennsylvania", "rhode island",
                    "south carolina", "tennessee", "texas", "vermont",
                    "virginia", "west virginia", "wisconsin")

choro <- CountyChoropleth$new(df_county)
choro$title <- "County in NMMAPS"
choro$set_zoom(eastern_states)

choro$ggplot_scale <- scale_fill_manual(name = "County status",
                                        values = c("white",
                                               "orange",
                                               "red"),
                                        labels = c("Not in NMMAPS",
                                               "In NMMAPS without Storm",
                                               "In NMMAPS with Storm"), 
                                        guide = guide_legend(reverse = TRUE))


choro$render()
```

[BA: The plot you get back from `choro$render()` is a ggplot object. You should be able to add on points based on latitude and longitude (this is how we're adding the hurricane tracks in our storm mapping). Here's a start for you. You'll want to remove all of the original NMMAPS locations that are not in the eastern states we're considering. If it's not too hard, you could also do the circles in orange or red for in / not in the storm]

```{r}
city.char <- readRDS("data/city.char.rds")

# remove locations not in the eastern states
city.char$statename <- tolower(city.char$statename)
latlong <- subset(city.char, city.char$statename %in% eastern_states)
latlong$lat <- as.numeric(latlong$lat)
latlong$long <- -1 * as.numeric(latlong$long)

# get value for colour
latlong <- left_join(latlong, get.city.county, by = "city")

a <- choro$render()
a + geom_point(data = latlong, aes(x = long, y = lat, group = NULL, 
                                   colour = factor(value)),
                                   shape = 1, size = 5, show.legend = FALSE) +
    scale_colour_manual(values = c("orange", "red")) 
ggsave("Plots/county_in_NMMAPS.png", width = 12, height = 9)
```


### Check out the map (based on 75mm rain exposure)

* St. Louis city (fips = 29510) in Missouri has no storm during our study period. 

* Storms hit the southeast part of TX state. Three counties don't have storms in the map.
    + El Paso county (48141, west corner of TX) is a low risk area. 
    + Travis county (48453) is a high risk hurricane zone. Travis county is not exposed to storm based on the selected values for rain and distance, and would be considered as storm county if we change the exposure condition.
    + Lubbock county (48303) is a low risk hurricane zone with only one hurricane recorded since 1930.
    + Our package seems not cover one big storm, "Lidia-1993" which hit Travis county and most part of TX. [BA: Hurricane Lidia in 1993 was a storm from the Pacific Ocean, rather than the Atlantic Ocean. We're restricting our analysis to the Atlantic basin storms, so that's why we don't have Lidia in our dataset.]

* Jefferson (21111) and Fayette (21067) are the two counties in Kentucky. Only Jefferson was exposed to storms in the map. The two counties might be at similar risk for hurricane. 
